Fast and Accurate Determination of Phase Transition Temperature in Computer Simulation

### 2018-12-19 manuscript20181218.tex
Revised in Sep 18,2018

\section{Results and Discussion} 
The direct coexistence technique is to put in contact two or more phases in one simulation box and study its thermodynamics. In our work, we still build our system in such classic way to clarify the effect induced by crystal lattice and interfacial crystal plane. And of course, save computational cost especially in freezing of full atom water. Water/ice coexisting systems are balanced at 250K for 100 ps, these short simulations are not enough for system to attain equilibrium, but able to relax the interfaces and a energy value $H=E+PV$ for NPT or $E$ for NVT ensemble can be extract to guide setting  energy value of GCE simulations\cite{Xu2012,Xu2015}. For mw ice Ih/water case, this value is around $-10.7 kcal/mol$  ,  the GCE equilibrated conformation with energy value range from $-9.9$ to $-11.5$ are shown in Fig\ref{fig:conformation}. Since water/ice transition involves no significant volume change, the energy value identifies the numerical average of water and ice energy, a lower energy value means more ice content. It is quite the same with unusal NPT ensemble if the equilibrated conformation is bulk phase, since energy distribution in bulk phase is always unimodal. That is, when H = -9.9,-10.1,-11.5 , GCE ensembles act like NPT ensemble of bulk water at 312K, 286K and bulk ice at 268K, as the evolutions of GCE ensembles  shown in Fig.\ref{fig:evolution}.
\begin{figure}[ht]
\centering{}\includegraphics[width=0.5\textwidth]{conf.png} 
\caption{The conformations of GCE ensembles at distinct energy levels: From bulk water above room temperature to completely freezed ice Ih.
\label{fig:conformation} }
\end{figure}

@@

This approach is also tested with full-atom water models, as shown in Fig. \ref{fig:PTtemp-2005} and Fig. \ref{fig:PTtemp-ICE}. In order to save expensive cpu cost, smaller MPS composed of 4584 water molecules are applied, and run for 40ns at most to collect trajectories. This time scale is not very sufficient for the MPS to attain equilibrium, especially in highly supercooled MPS case, since the growth rate is proportional to the diffusivity of supercooled liquid water\cite{Tanaka2003}. However, the results still outline a rough but explicit picture of this GCE method. Once the MPS is built, simulations covering the energy space of multiple-phase coexisting state will spontaneously reveal the phase transitional temperature. For TIP4P-2005 water model, the result is $250.0\pm0.6K$, match with $T_m(DC)=249\pm3K$, $T_m(FE)=252\pm5K$ and $T_m(DC)=249.5\pm0.1K$ from previous research\cite{Conde2013,Conde2017}, as Fig. \ref{fig:PTtemp-2005} shows. The energy evolutions are also presented in the picture. Temperature of three MPS with $H=-4.25,-4.16,-4.06 kcal/mol$ come to a same level within 20ns, further 20ns run are applied to collect data of equilibrated systems. The TIP4P-ICE water model have a better performance in generating ice/water phase diagram. As shown in Fig. \ref{fig:PTtemp-ICE}, our result $T_m=270.2\pm0.3K$ shows an excellent consistency with documents\cite{Conde2017,Abascal2005}.
 \begin{figure}[ht]
\centering{}\includegraphics[width=0.5\textwidth]{PTtemp-2005-insec.png} 
\caption{GCE method determined phase transition temperature for TIP4P-2005 water model. cyan region refers to $252\pm5K$ from free energy route, yellow region refers to $249\pm3K$ of direct coexistence route\cite{Conde2013} . 
\label{fig:PTtemp-2005}} 
\end{figure}

 \begin{figure}[ht]
\centering{}\includegraphics[width=0.5\textwidth]{PTtemp-ICE-insec.png} 
\caption{GCE method determined phase transition temperature for TIP4P-ICE water model. The melting point(272.2K) proposed in previous work\cite{Abascal2005} is plotted as dash line. 
\label{fig:PTtemp-ICE}} 
\end{figure}

@@

 \begin{figure}[ht]
\centering{}\includegraphics[width=0.5\textwidth]{PTtemp-lj.png} 
\caption{GCE method determined phase transition temperature $T_m=0.619\pm0.003\epsilon/k_B$for LJ model.
\label{fig:PTtemp-lj}} 
\end{figure}
A further universality test has been carried out in LJ model. As a well studied simple model in molecular dynamics simulation, The triple point temperature of LJ particles is known as $T^*=0.617\epsilon/k_B$, showing excellent agreement with our result  $T_m=0.619\pm0.003\epsilon/k_B$. The melting temperature shows a same trend in MPS of water models, decreases while the solid phase is getting thinner.

>>>>

\section{Results and Discussion} 
The direct coexistence technique is to put in contact two or more phases in one simulation box and study its thermodynamics. In our work, we still build our system in such classic way to clarify the effect induced by crystal lattice and interfacial crystal plane. And of course, save computational cost especially in freezing of full atom water. Water/ice coexisting systems are balanced at 250K for 100 ps, these short simulations are not enough for system to attain equilibrium, but able to relax the interfaces and a energy value $H=E+PV$ for NPT or $E$ for NVT ensemble can be extract to guide setting  energy value of GCE simulations\cite{Xu2012,Xu2015}. For mW ice Ih/water case, this value is around $-10.7 kcal/mol$  ,  the GCE equilibrated conformation with energy value range from $-9.9$ to $-11.5$ are shown in Fig\ref{fig:conformation}. Since water/ice transition involves no significant volume change, the energy value identifies the numerical average of water and ice energy, a lower energy value means more ice content. It is quite the same with unusal NPT ensemble if the equilibrated conformation is bulk phase, since energy distribution in bulk phase is always unimodal. That is, when H = -9.9,-10.1,-11.5 , GCE ensembles act like NPT ensemble of bulk water at 312K, 286K and bulk ice at 268K, as the evolutions of GCE ensembles  shown in Fig.\ref{fig:evolution}.
\begin{figure}[ht]
\centering{}\includegraphics[width=0.5\textwidth]{conf.png} 
\caption{The conformations of GCE ensembles at distinct energy levels: From bulk water above room temperature to completely freezed ice Ih.
\label{fig:conformation} }
\end{figure}

@@

This approach is also tested with full-atom water models and LJ models as shown in Table. \ref{tab:tab1}. In order to save expensive cpu cost, smaller MPS composed of 4584 water molecules are applied in full-atom water systems, and run for 40ns at most to collect trajectories. This time scale is not very sufficient for the MPS to attain equilibrium, especially in highly supercooled MPS case, since the growth rate is proportional to the diffusivity of supercooled liquid water\cite{Tanaka2003}. However, the results still outline a rough but explicit picture of this GCE method. Once the MPS is built, simulations covering the energy space of multiple-phase coexisting state will spontaneously reveal the phase transitional temperature. For TIP4P-2005 water model, the result is $250.0\pm0.6K$, match with $T_m(DC)=249\pm3K$, $T_m(FE)=252\pm5K$ and $T_m(DC)=249.5\pm0.1K$ from previous research\cite{Conde2013,Conde2017}. The TIP4P-ICE water model have a better performance in generating ice/water phase diagram, our result $T_m=270.2\pm0.3K$ shows an excellent consistency with documents\cite{Conde2017,Abascal2005}. As a well studied simple model in molecular dynamics simulation, The triple point temperature of LJ particles is known as $T^*=0.617\epsilon/k_B$, showing excellent agreement with our result  $T_m=0.619\pm0.003\epsilon/k_B$.

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\begin{table}
\caption{Melting points Tm of the mW, TIP4P/2005, TIP4P/Ice water and LJ models as obtained from different methodologies (free energy calculations, direct coexistence technique and grand-canonical ensemble approach).}

\centering{}%
\begin{tabular}{ccccc}
%\toprule 
\hline
{ Method} & {mW} & {TIP4P-2005}  & {TIP4P-ICE}  & {LJ} \tabularnewline
%\midrule
\hline
{ FE} & { $274.6\pm1$\cite{Molinero2009}} & {$252\pm6$\cite{Abascal2005}}  & {$272\pm6$\cite{Abascal2005}}  & {0.617\cite{Broughton1986}} \tabularnewline
\hline
{ DC} & { - } & \tabincell{c}{$249\pm3$\cite{Conde2013}\\$249.5\pm0.1$\cite{Conde2017}} & \tabincell{c}{$268\pm2$\cite{GarciaFernandez2006}\\$269.8\pm0.1$\cite{Conde2017}}  & {-}  \tabularnewline
\hline
{ GCE} & { $275.1\pm0.3$} & {$250.0\pm0.6$ }  &{$270.2\pm0.3$} & {$0.619\pm0.003$} \tabularnewline
\hline
\end{tabular}
\label{table:tab1}
\end{table}

@@
